Appendix B: Supplemental Results

### 2020-02-07 ms/supp/App_1.Rmd
slope comparisons updated

\section{General model structure}
\subsection{Integral Projection Model overview}
To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions. The equations below apply to both models.

IPMs use the size distribution, $z$, of individuals at time $t$, along with a kernel $K(z^{\prime}, z)$, to predict the size distribution, $z^{\prime}$, at time $t+1$:
\begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

@@

\newpage
\section{Regression equations for the virtual species}
We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species. Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$.

@@

@@

>>>>

\section{General model structure}
\subsection{Integral Projection Models (IPMs)}
To generate fully-known true distributions for the virtual species, we used the general structure of an Integral Projection Model (IPM) to calculate the intrinsic growth rate $\lambda$ in each cell of the gridded landscape, and adapted the regression-based structure of an IPM into an individual-level, simulation-based cellular automata (CA) model to produce spatiotemporally dynamic abundance distributions.

IPMs (Easterling 2000; Merow et al. 2014, 2017) use the size distribution, $z$, of individuals at time $t$, along with a kernel $K(z^{\prime}, z)$, to predict the size distribution, $z^{\prime}$, at time $t+1$:
\begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

@@

\newpage
\section{Regression equations for the virtual species}
We used the above IPM structure, including a seed bank and dispersal, as the basis for the virtual species. Thus, the population in each cell $i$ has size distribution $z$, where $z^{\prime}$ is the size distribution the next year, and $\boldsymbol{z_i}$ is a matrix with columns of 1 (intercept) and $z$. For the IPMs, we used the following regression equations to populate discretized transition matrices for each cell to then calculate the deterministic intrinsic growth rate $\lambda$. For the CA\textsubscript{i} models, we used the same regression equations to generate expected values for each individual, and then drew simulated outcomes from the specified distribtutions.

@@

Easterling, M. R., Ellner, S. P., & Dixon, P. M. (2000). Size-specific sensitivity: Applying a new structured population model. Ecology, 81(3), 694–708. https://doi.org/10.1890/0012-9658(2000)081[0694:SSSAAN]2.0.CO;2

@@

Merow, C., Dahlgren, J. P., Metcalf, C. J. E., Childs, D. Z., Evans, M. E. K., Jongejans, E., … Mcmahon, S. M. (2014). Advancing population ecology with integral projection models: A practical guide. Methods in Ecology and Evolution, 5(2), 99–110. https://doi.org/10.1111/2041-210X.12146

Merow, C., Bois, S. T., Allen, J. M., Xie, Y., & Silander, J. A. (2017). Climate change both facilitates and inhibits invasive plant ranges in New England. Proceedings of the National Academy of Sciences, 114(16), E3276–E3284. https://doi.org/10.1073/pnas.1609633114

### 2020-02-07 ms/supp/App_2.Rmd
slope comparisons updated

\newpage
\section{Estimated slopes}
The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared. The pattern of accuracy across scenarios reflects the observed trends in TSS. The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub. The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older. This effect is absent from the biennial regressions, as individuals are not long-lived.
\begin{figure}
	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_shrub.pdf}
	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub. Regressions are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_biennial.pdf}
	\caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.}
\end{figure}

>>>>

\newpage
\section{Estimated slopes}
The IPM and CA\textsubscript{i} models fit regressions of the same form as the generative models, and so error in the estimated slopes can be directly compared. The pattern of accuracy across scenarios reflects the observed trends in TSS. The effect of data scenario is minimal, with the exception of non-equilibrium for the shrub. The slopes clearly show the improvement from sampling newer populations with a broader age distribution, relative to all other scenarios where populations tended to be older. This effect is absent from the biennial regressions, as individuals are not long-lived.
\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
	\caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the vital rate regressions for the shrub. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. For (a) survival probability, the slopes are estimated more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. In contrast, (b) growth, (c) flowering probability, and (e) germination probability show little variation among scenarios, while for (d) seed production, added \emph{measurement error} increased variability. Note that the axes vary among panels.}
\end{figure}

\begin{figure}
	\centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
	\caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the vital rate regressions for the biennial. Each density curve shows the distribution of fitted slope values across 100 data sets used in the IPM and CA\textsubscript{i} models. The true value is indicated by the vertical dashed line. The fitted slopes were consistent with minimal differences among data scenarios for (a) survival probability, (b) growth, (c) flowering probability, (d) seed production, and (e) germination probability. Peaks at 0 represent data sets where the optimal model did not include that term. Note that the axes vary among panels.}
\end{figure}

### 2020-02-07 ms/supp/App_A.tex
slope comparisons updated

To generate fully-known true distributions for the virtual species, we
used the general structure of an Integral Projection Model (IPM) to
calculate the intrinsic growth rate \(\lambda\) in each cell of the
gridded landscape, and adapted the regression-based structure of an IPM
into an individual-level, simulation-based cellular automata (CA) model
to produce spatiotemporally dynamic abundance distributions. The
equations below apply to both models.

IPMs use the size distribution, \(z\), of individuals at time \(t\),
along with a kernel \(K(z^{\prime}, z)\), to predict the size
distribution, \(z^{\prime}\), at time \(t+1\): \begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

@@

We used the above IPM structure, including a seed bank and dispersal, as
the basis for the virtual species. Thus, the population in each cell
\(i\) has size distribution \(z\), where \(z^{\prime}\) is the size
distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
columns of 1 (intercept) and \(z\).

@@

@@

>>>>

To generate fully-known true distributions for the virtual species, we
used the general structure of an Integral Projection Model (IPM) to
calculate the intrinsic growth rate \(\lambda\) in each cell of the
gridded landscape, and adapted the regression-based structure of an IPM
into an individual-level, simulation-based cellular automata (CA) model
to produce spatiotemporally dynamic abundance distributions.

IPMs (Easterling 2000; Merow et al.~2014, 2017) use the size
distribution, \(z\), of individuals at time \(t\), along with a kernel
\(K(z^{\prime}, z)\), to predict the size distribution, \(z^{\prime}\),
at time \(t+1\): \begin{equation}
n_{t+1}(z^{\prime}) = \int_{\Omega} K(z^{\prime}, z) n_t(z) dz
\end{equation}

@@

We used the above IPM structure, including a seed bank and dispersal, as
the basis for the virtual species. Thus, the population in each cell
\(i\) has size distribution \(z\), where \(z^{\prime}\) is the size
distribution the next year, and \(\boldsymbol{z_i}\) is a matrix with
columns of 1 (intercept) and \(z\). For the IPMs, we used the following
regression equations to populate discretized transition matrices for
each cell to then calculate the deterministic intrinsic growth rate
\(\lambda\). For the CA\textsubscript{i} models, we used the same
regression equations to generate expected values for each individual,
and then drew simulated outcomes from the specified distribtutions.

@@

Easterling, M. R., Ellner, S. P., \& Dixon, P. M. (2000). Size-specific
sensitivity: Applying a new structured population model. Ecology, 81(3),
694--708.
\url{https://doi.org/10.1890/0012-9658(2000)081\%5B0694:SSSAAN\%5D2.0.CO;2}

@@

Merow, C., Dahlgren, J. P., Metcalf, C. J. E., Childs, D. Z., Evans, M.
E. K., Jongejans, E., \ldots{} Mcmahon, S. M. (2014). Advancing
population ecology with integral projection models: A practical guide.
Methods in Ecology and Evolution, 5(2), 99--110.
\url{https://doi.org/10.1111/2041-210X.12146}

Merow, C., Bois, S. T., Allen, J. M., Xie, Y., \& Silander, J. A.
(2017). Climate change both facilitates and inhibits invasive plant
ranges in New England. Proceedings of the National Academy of Sciences,
114(16), E3276--E3284. \url{https://doi.org/10.1073/pnas.1609633114}

### 2020-02-07 ms/supp/App_B.tex
slope comparisons updated

This appendix contains supplemental tables and figures to provide
additional detail for the results, including maps of \(\lambda\) and
\(N\) for each species (true, predicted, difference), figures and tables
for sensitivity and specificity, and additional detail on predicted
slopes for the IPM and CA\textsubscript{i} models.

\section{Prediction maps}

@@

\newpage
\begin{figure}
    \centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Sens_mn+CI.jpg}
    \caption{\label{fig:SensitivityMed} Sensitivity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$. Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal). Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

\newpage
\begin{figure}
    \centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Spec_mn+CI.jpg}
    \caption{\label{fig:SpecificityMed} Specificity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$. Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal). Specificity represents the proportion of true absences that were correctly predicted as absences.}
\end{figure}

@@

\newpage
\begin{figure}
    \centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
    \caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the \emph{ideal} scenario. Specificity represents the proportion of true absences that were correctly predicted as absences.}
\end{figure}

@@

\begin{figure}
    \centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_shrub.pdf}
    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub. Regressions are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution.}
\end{figure}

\begin{figure}
    \centering\includegraphics[width=.8\linewidth]{../../figs/diag/slope_box_biennial.pdf}
    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios.}
\end{figure}

>>>>

This appendix contains supplemental figures to provide additional detail
for the results, including figures for sensitivity and specificity, and
additional detail on predicted slopes for the IPM and
CA\textsubscript{i} models.

@@

\begin{figure}
    \centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Sens_mn+CI.jpg}
    \caption{\label{fig:SensitivityMed} Sensitivity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$. Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal). Sensitivity represents the proportion of true presences that were correctly predicted as presences.}
\end{figure}

\begin{figure}
    \centering\includegraphics[width=.75\linewidth]{../../figs/Supp_Spec_mn+CI.jpg}
    \caption{\label{fig:SpecificityMed} Specificity mean and 95\% confidence intervals across 100 sampled data sets for each SDM and scenario, compared to true distributions defined by $\lambda > 1$ and $N > 0$. Scenarios include: no sampling or modeling issues (ideal), sampling issues (measurement error, sampling bias, non-equilibrium), and modeling issues (incorrect covariates, no seed bank, under dispersal, over dispersal). Specificity represents the proportion of true absences that were correctly predicted as absences.}
\end{figure}

@@

\begin{figure}
    \centering\includegraphics[width=.9\linewidth]{../../figs/Supp_SpecvIdeal.jpg}
    \caption{\label{fig:SpecificityvIdeal} Effect of scenario on median specificity relative to the \emph{ideal} scenario. Specificity represents the proportion of true absences that were correctly predicted as absences.}
\end{figure}

@@

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_shrub.pdf}
    \caption{\label{fig:slopesShrub} Effect of data scenario on slope estimates for the shrub. Survival regressions for the shrub are parameterized more accurately when using \emph{non-equilibrium} populations which contain a more even size distribution. Note that the axes vary among panels.}
\end{figure}

\begin{figure}
    \centering\includegraphics[width=\linewidth]{../../figs/diag/slope_dens_biennial.pdf}
    \caption{\label{fig:slopesBiennial} Effect of data scenario on slope estimates for the biennial Regressions are parameterized consistently with minimal differences among data scenarios. Peaks at 0 represent data sets where the optimal model did not include that term. Note that the axes vary among panels.}
\end{figure}
